Acceleration of plague outbreaks in the second pandemic

### 2020-07-12 supp.Rnw
OK except for supp_crossrefs.tex, which I can no longer include in ms.tex and consequently refs to supp are broken in ms

\documentclass[9pt,twocolumn,twoside,lineno,table]{pnas-new}
% Use the lineno option to display guide line numbers if required.
%% knitr requires documentclass to be the first line of the file
%% Created: 19 Mar 2017
%% Changed: 22 Apr 2017
%%          30 Apr 2017

@@

%
\begin{table}%[h!]
\caption[Population estimates for London]{Population estimates for London previously published by Finlay \cite{Finl81,FinlShea86}\cite[p.\,108]{Kryl11} and Creighton \cite[p.\,660]{Crei65}.  These data are plotted in \cref{F:london_pop}.}
\label{tab:pop}
\begin{center}
%% \input{analysis/tables/supppop.tex} % generated by analysis/tables/supppop.R
%% DE: I'm building the (somewhat different) table myself.
\renewcommand{\arraystretch}{1.0}
\begin{tabular}{c|r|l}
%%\hline
{\bfseries Year} &
{\bfseries Population} &
{\bfseries Source}\\
\hline
  1340 &  45,000 & Finlay \\ 
  1500 &  50,000 & Finlay \\ 
  1550 & 110,000 & Finlay \\ 
  1600 & 180,000 & Finlay \\ 
  1603 & 250,000 & Creighton \\
  1625 & 320,000 & Creighton \\
  1650 & 340,000 & Finlay \\ 
  1665 & 460,000 & Creighton \\
  1700 & 445,000 & Finlay \\ 
\end{tabular}
\end{center}
\label{T:london_pop}
\end{table}
%

%%\FloatBarrier
\section*{Disease-related data sources}

@@

%%\newcommand{\mywidth}{0.8\textwidth}
\newcommand{\mywidth}{\textwidth}
%
\begin{figure*}%[h!]
\begin{center}
\includegraphics[width=\mywidth]{analysis/plots/plotallscaled.crop.pdf}
\end{center}
\caption[Plague deaths per 1000 per week]{Plague deaths per 1000 persons per week, from 1563 to 1666.
Epidemics that exceeded 5 on this scale were classified as \emph{major}.
Analysis of the major epidemics is shown in \cref{tab:params,F:timeseries,F:combdata}.  Corresponding analyses of all of the minor epidemics are presented
in \cref{tab:minorparams,F:seasonalgrowth}.
}
\label{F:all1600s.scaled}
\end{figure*}
%

%%\clearpage
\section*{Generation interval for bubonic plague}\label{sec:gentime}

@@

In the idealized situation in which the generation interval is exponentially distributed (as in the standard SIR model) then \cite{WallLips07}, as mentioned in the Discussion in the main text, 
%
\begin{linenomath*}
\begin{equation}\label{E:rofR}
r = \frac{\R_0-1}\gentime \,,
\end{equation}
\end{linenomath*}
%
where $\gentime$ is the mean generation interval.  Consequently, if this simple relationship holds and the mean generation interval \emph{does not change}, then a change in growth rate $r_1\to r_2$ implies a change in basic reproduction number $\R_{0,1}\to\R_{0,2}$, where
%
\begin{linenomath*}
\begin{equation}\label{E:1to2}
\frac{\R_{0,2}}{\R_{0,1}} = \frac{r_2\gentime+1}{r_1\gentime+1} \,.
\end{equation}
\end{linenomath*}
%
If we assume the estimate of $\gentime=18$ days obtained in
\cref{sec:gentime} above, and the MLE growth rates from early and late wills
listed in \cref{tab:params}, the relative change in reproduction
number that needs explanation is
<<rat_R0_calc,echo=FALSE>>=
ratR0 <- function(r){ (r / 365 * 18) + 1 }
load("analysis/fits/epochsum.RData")
r1 <- rvals["fit","early"]
r2 <- rvals["fit","late"]
R01 <- ratR0(r1)
R02 <- ratR0(r2)
R0ratio <- R02/R01
@ 
%
\begin{linenomath*}
\begin{equation}\label{E:Rratio}
\frac{\R_{0,2}}{\R_{0,1}} 
\approx
\frac
{\Sexpr{round(r2,1)} \cdot \frac{18}{365} + 1}
{ \Sexpr{round(r1,2)} \cdot \frac{18}{365} + 1}
\approx
\frac{\Sexpr{round(R02,2)}}{\Sexpr{round(R01,2)}} 
\approx \Sexpr{round(R0ratio,2)}
\,.
\end{equation}
\end{linenomath*}
%
where the second subscript (1 or 2) denotes the early or late epoch, respectively.

@@

\FloatBarrier
\clearpage
\section*{Supplementary tables cited in the main text}

@@

\FloatBarrier
\section*{Supplementary figure cited in the main text}

>>>>

\documentclass[9pt,twocolumn,twoside,lineno,table]{pnas-new}
% Use the lineno option to display guide line numbers if required.

%% knitr requires documentclass to be the first line of the file

@@

\section*{Disease-related data sources}

@@

\hypertarget{sec:gentime}{}
\section*{Generation interval for bubonic plague}

@@

In the idealized situation in which the generation interval is exponentially distributed (as in the standard SIR model) then \cite{WallLips07}, as mentioned in the Discussion in the main text, 
%
\begin{linenomath*}
\begin{equation}\label{E:rofR}
r = \frac{\R_0-1}\gentime \,,
\end{equation}
\end{linenomath*}
%
where $\gentime$ is the mean generation interval.  Consequently, if this simple relationship holds and the mean generation interval \emph{does not change}, then a change in growth rate $r_1\to r_2$ implies a change in basic reproduction number $\R_{0,1}\to\R_{0,2}$, where
%
\begin{linenomath*}
\begin{equation}\label{E:1to2}
\frac{\R_{0,2}}{\R_{0,1}} = \frac{r_2\gentime+1}{r_1\gentime+1} \,.
\end{equation}
\end{linenomath*}
%
If we assume the estimate of $\gentime=18$ days obtained
\hyperlink{sec:gentime}{above}, and the MLE growth rates from early and late wills
listed in \cref{tab:params}, the relative change in reproduction
number that needs explanation is
<<rat_R0_calc,echo=FALSE>>=
ratR0 <- function(r){ (r / 365 * 18) + 1 }
load("analysis/fits/epochsum.RData")
r1 <- rvals["fit","early"]
r2 <- rvals["fit","late"]
R01 <- ratR0(r1)
R02 <- ratR0(r2)
R0ratio <- R02/R01
@ 
%
\begin{linenomath*}
\begin{equation}\label{E:Rratio}
\frac{\R_{0,2}}{\R_{0,1}} 
\approx
\frac
{\Sexpr{round(r2,1)} \cdot \frac{18}{365} + 1}
{ \Sexpr{round(r1,2)} \cdot \frac{18}{365} + 1}
\approx
\frac{\Sexpr{round(R02,2)}}{\Sexpr{round(R01,2)}} 
\approx \Sexpr{round(R0ratio,2)}
\,.
\end{equation}
\end{linenomath*}
%
where the second subscript (1 or 2) denotes the early or late epoch, respectively.

@@

\section*{Supplementary tables cited in the main text}

\cref{tab:summarystats,tab:minorparams,tab:windows,tab:windowsminor}.

\section*{Supplementary figure cited in the main text}

\cref{F:seasonalgrowth}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%
%% FIGURES %%
%%%%%%%%%%%%%

\newcommand{\mywidth}{\textwidth}
%
\begin{figure*}%[h!]
\begin{center}
\includegraphics[width=\mywidth]{analysis/plots/plotallscaled.crop.pdf}
\end{center}
\caption[Plague deaths per 1000 per week]{Plague deaths per 1000 persons per week, from 1563 to 1666.
Epidemics that exceeded 5 on this scale were classified as \emph{major}.
Analysis of the major epidemics is shown in \cref{tab:params,F:timeseries,F:combdata}.  Corresponding analyses of all of the minor epidemics are presented
in \cref{tab:minorparams,F:seasonalgrowth}.
}
\label{F:all1600s.scaled}
\end{figure*}
%

{%%%%%%%%%%%%
%% TABLES %%
%%%%%%%%%%%%

%
\begin{table*}%[h!]
\caption[Population estimates for London]{Population estimates for London previously published by Finlay \cite{Finl81,FinlShea86}\cite[p.\,108]{Kryl11} and Creighton \cite[p.\,660]{Crei65}.  These data are plotted in \cref{F:london_pop}.}
\label{tab:pop}
\begin{center}
%% \input{analysis/tables/supppop.tex} % generated by analysis/tables/supppop.R
%% DE: I'm building the (somewhat different) table myself.
\renewcommand{\arraystretch}{1.0}
\begin{tabular}{c|r|l}
%%\hline
{\bfseries Year} &
{\bfseries Population} &
{\bfseries Source}\\
\hline
  1340 &  45,000 & Finlay \\ 
  1500 &  50,000 & Finlay \\ 
  1550 & 110,000 & Finlay \\ 
  1600 & 180,000 & Finlay \\ 
  1603 & 250,000 & Creighton \\
  1625 & 320,000 & Creighton \\
  1650 & 340,000 & Finlay \\ 
  1665 & 460,000 & Creighton \\
  1700 & 445,000 & Finlay \\ 
\end{tabular}
\end{center}
\label{T:london_pop}
\end{table*}
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
